
\documentclass{article}

\usepackage[usenames,dvipsnames]{xcolor}


\usepackage{listings}
\usepackage{multicol}


\lstset{
    frame=tb, % draw a frame at the top and bottom of the code block
    tabsize=4, % tab space width
    showstringspaces=false, % don't mark spaces in strings
    numbers=left, % display line numbers on the left
    commentstyle=\color{green}, % comment color
    keywordstyle=\color{blue}, % keyword color
    stringstyle=\color{red}, % string color
    identifierstyle=\color{grey},
}


\definecolor{diffstart}{named}{gray}
\definecolor{diffincl}{named}{OliveGreen}
\definecolor{diffrem}{named}{red}

\lstdefinelanguage{diff}{
    morecomment=[f][\color{diffstart}]{@@},
    morecomment=[f][\color{diffincl}]{+},
    morecomment=[f][\color{diffrem}]{-},
  }

\input{listings-coq}

\begin{document}

\title{Verifying existing ASN.1 compiler}

\author{Nika Pona, Vadim Zaliva \\
        Digamma.ai}

\maketitle

\begin{abstract}

\end{abstract}
\begin{enumerate}

\item Intro
  \begin{itemize}
  \item ASN.1 intro and example 
  \item Mention some common problems with ASN.1 parsers
    \item X.509 subset of ASN.1
  \end{itemize}
\item Approaches to verification of parsers
  \begin{itemize}
  \item Functional parsers (Narcissus)
  \item Extraction to C, partial correctness (Everparse, Galois)
   \item Verifying existing C code vs extraction
  \end{itemize}
  \item asn1c compiler in a nutshell
  \item Example of strtoimax of asn1c
\begin{itemize}
\item Bugs discovered
  \item Op. semantics and VST proofs
  \item Lessons learned + plans
  \end{itemize}
\item Current work
  \begin{itemize}
  \item exploit FSM-like organization of the parser
    \end{itemize}

\end{enumerate}

\section{Intro}
 ASN.1 intro and simple example 

 The Computer Vulnerabilities and Exposures (CVE) database [7a] lists critical ASN.1-related bugs found each year in the existing systems, and there have already been noteworthy exposures [8] that although not as dire to security as first feared [9] clearly spell out a first awareness of the vast risk and exposure. We analyzed the last 4 years of ASN.1-related issues reported in Computer Vulnerabilities and Exposures (CVE) database [7b]. Among vulnerabilities studied were CVEs for various software and hardware products and vendors, including Apple, axTLS, Botan, Bounty Castle, librcrypto++, libtasn, LibTomCrypt, Linux Kernel, MatrixSSL, Mozilla NSS (Firefox), Objective Systems, OpenSSL, PolarSSL, RSA BSAFE, Samba, Samsung, Snapdragon, strongSwan, and Wireshark. It was found that 39 out of 52 problems analyzed were related to memory safety, 6 related to stack and heap bounds checking, and 3 related to issues caused by applications accepting not well-formed ASN.1 input. Having proved just six formal properties would have prevented 49 out of 52 vulnerabilities, that is, more than 90\% of reported vulnerabilities.

 \section{Other approaches to verification of parsers}

  Functional parsers (Narcissus)

  Extraction to C (Everparse, Galois)

  Project Everest [19] is our most direct thematic competition, and some of the high-level Project Everest documentation makes passing mention to ASN.1. The Project Everest stewards adopted a different approach based on F* [4], and the Project Everest ASN.1 work  appears at best still in far distant plans.

It is also noted that Galois did some work on ASN.1 verification in the past (circa 2012) [11]. It appears that Galois abandoned [12] the goal of full ASN.1 verification that we pursue in our project with our more pragmatic approach; Galois is now only exploring a limited subset ASN.1 verification adequate for the “vehicle-to-vehicle” (V2V) market [13], but that particular subset has limited broader applicability and the Galois effort appears encumbered by aspects of an unsuccessful  approach. It is noted that although our goal is eventually to verify all ASN.1, we decided to start with a different (X.509-related) subset of ASN.1; this subset is reasonably small, but X.509 is so widely used that our initial verified implementation of our chosen subset will have a large volume and wide range of immediate commercial applications.

\subsection{Our approach}
  Our technology and approach will provide an entirely new level of software and protocol verification employing recently-emergent provable full-functional correctness methods.
  Systems and methods to date either (a) automatically tested but not formally verified (b) use verification approach which rely on automatic extraction from executable specifications (for example involving network stack synthesis [31], optimizing compilers [3], cryptographic libraries [6], and encoder/decoders [32]), or (c) apply a form of formal verification which only proves partial correctness properties (partial verification of NAT stack only proving parts of DPDK are specification compliant [33], partial verification of Linux kernel TCP implementation with 55\% line coverage and 92\% protocol coverage [34]). Consequently, (a) and (c) do not provide sufficient correctness guarantees, while (c) is often impractical due to poor performance and compatibility limitations. In contrast, we pursue a far deeper and comprehensive verification approach to performance and portability and seek to prove actual industrial-level C-code implementation.

  Our plan:
  \begin{itemize}

  \item Formalize the X.509 part of the ASN.1 standard.  
   
  \item Refactor ASN1C code to make it suitable for verification.

   \item Establish correctness of encoders/decoders for primitive types.

  \item Establish correctness of encoders/decoders for constructed types.

  \item Prove high-level properties.

   \item Experiment with OCaml code extraction from Executable Specifications.
  \end{itemize}

\section{Veryfing a function from \texttt{asn1c} compiler}

  At \url{Digamma.ai} we are working on formal verification of existing {\bf imperative} programs using Coq. We took a function \texttt{asn\_strtoimax\_lim} from \texttt{asn1c} compiler to test several approaches on a simple real-life example. The function is relatively simple, but at the same time uses many features of C that make verifying imperative programs challenging. XER decoding functions for INTEGER, OBJECT-IDENTIFIER and RELATIVE-OID types critically depend on this function. XER decoding is pretty straightforward\footnote{INTEGERs are encoded in text as their decimal representation, OBJECT IDENTIFIERs are encoded as decimal numbers in the order of the object identifier's components.}, however, we found that it is compromised by three bugs in this function.
    Informal specification from the comments: 
  \begin{quote}

 { \it Parse the number in the given string until the given *end position,
 returning the position after the last parsed character back using the
 same (*end) pointer.
 WARNING: This behavior is different from the standard strtol/strtoimax(3). }
\end{quote}
  
  {\fontsize{8}{4}\selectfont  \lstinputlisting[language=C]{asn_strtoimax_lim_old.c}}
  \skip
  
\paragraph{Negative range bug}

When we go beyond allowed \textit{int} range, a wrong result is given for some inputs\footnote{Assume we are working on a 8-bit system and maximal signed int \texttt{MAX\_INT} is 127}:

\skip
\begin{tabular}{|l|l|}
 \hline
input & \texttt{-128} \\
intmax & \texttt{\ 127}\\
upper boundary& \texttt{\ 12} \\
last digit max& \texttt{\ 7}\\
return & \color{green}\texttt{\ ASN\_STRTOX\_ERROR\_RANGE}\\
\hline 
input & \texttt{-1281} \\
intmax & \texttt{\ 127}\\
upper boundary& \texttt{\ 12} \\
last digit max& \texttt{\ 7}\\
return & \color{red}\texttt{\ -127, ASN\_STRTOX\_OK}\\
\hline
    \end{tabular}

 \skip
 
This happens whenever the input string represents a number smaller than \texttt{MIN\_INT}, due to the fact that absolute value of \texttt{MIN\_INT} is greater than \texttt{MAX\_INT}, thus negative number cannot be treated as $\mathrm{value}\times\mathrm{sign}$ when $\mathrm{value}$ is represented as \textit{int}. The bug  was filed and promptly fixed by developers:

   
  
  {\fontsize{8}{4}\selectfont \lstinputlisting[language=diff]{bug1fix.diff}}

  
  \paragraph{Memory store bug}
  We uncovered a curious bug in the \emph{fixed} version. Note that we store \texttt{str} before reading from \texttt{str}: 
     \begin{lstlisting}[language=C]
  if(str < *end) {
    *end = str;
    if(*str >= 0x30 && *str <= 0x39){
      return ASN_STRTOX_ERROR_RANGE;
    } else {
      *intp = sign * value;
      return ASN_STRTOX_EXTRA_DATA;
    }
  }
  \end{lstlisting}
  
     We can construct an example where we get wrong results by overwriting data in \texttt{str}. Let minimal signed int \texttt{MIN\_INT} = $-4775808$ and \texttt{*str} = {\color{red}2d 34 37 37 35 38 30 31 31 31} (stands for ``-477580111'').
     \newpage

\begin{multicols}{2}
  
    {\bf Scenario 1}:

    Assume that \texttt{*end = str + 9} and

    \texttt{end} $\geq$ $\texttt{str + 9}$.

    {\color{red}2d 34 37 37 35 38 30 {\color{blue}$\overbrace{31}^{\texttt{str + 7}}$} 31 31

      $\ldots$ $\overbrace{\texttt{X}}^{\texttt{end}}$}

    Then at \texttt{str + 7} we store

    \texttt{*end = (str + 7)}

     Let \texttt{str + 7 = 21 21 21 26} 

     {\color{red}2d 34 37 37 35 38 30 31 {\color{blue} $\overbrace{31}^{\texttt{str + 8}}$} 31

       $\ldots$ $\overbrace{\texttt{21 21 21 26}}^{\texttt{end}}$}
     
     
     And since at \texttt{str + 8} we read `1'
     

     
     The output is  {\color{green}\texttt{ASN\_ERROR\_RANGE}}
     
     \columnbreak
     
    {\bf Scenario 2}:

    Assume that \texttt{*end = str + 9} and

    \texttt{end = str + 7}:
    
    {\color{red}2d 34 37 37 35 38 {\color{blue} $\overbrace{30 \; 31 \; 31 \; 31}^{\texttt{end}}$}}
    

    Then at \texttt{str + 7} we store

    \texttt{*end = str + 7}

     Let \texttt{str + 7 = 21 21 21 26} 

   
    {\color{red}2d 34 37 37 35 38 30 {\color{blue}$\overbrace{21 \; 21 \; 21 \; 26}^{\texttt{end}}$}}

    (stands for $``-477580!!!\&''$)

    And since at \texttt{str + 8} we read `!' 

    The output is {\color{red}\texttt{ASN\_EXTRA\_DATA}}
    \end{multicols}
    

Hence, when the value of the \texttt{end} pointer is treated as a part of the input data, there is a bug where the resulting error value could be incorrect. On the other hand, it is hard to think of a legitimate use-case where the pointer would be a part the input data. Under such interpretation, there is an implicit pre-condition in the specification, mandating that:

{\small
\texttt{(*end < end) || (end + sizeof(const char *) <= str)}}


\paragraph{Specification ``bug''}
After addressing the two bugs we discovered we were able to successfully verify that the function finally corresponds to the specification we wrote for. However, it was noticed the following behavior:

For input \texttt{``a''} it stores value 0 and returns {\color{green}\texttt{ASN\_STRTOX\_EXTRA\_DATA}} (same behaviour as on input \texttt{``0a''}), which could be unintended by authors. \\

This code was part of the library for 15 years. The library is covered by extensive unit and randomized tests. It is used in production by multiple users. Yet, the vulnerabilities are there and pose potential problems.

\section{Our approach}


\subsection{Machine arithmetic}

The first bug is related to data type ranges and modulo integer arithmetic. These sort of problems are fairly common and require careful coding to be avoided. Formal verification enforces a strict mathematical model of all computer arithmetic and invariably exposes all such bugs. We use CompCert's \texttt{Integer} library that provides theory of 8-, 16-, 32- and 64-bit integers and 32- and 64-bit pointers. It has some basic lemmas, however, doing proofs by hand is quite tedious, however automation can be achieved here for programs that allow no overflow, since then one can use \texttt{Micromega}\footnote{\url{https://coq.inria.fr/refman/addendum/micromega.html#micromega-tactics-for-solving-arithmetic-goals-over-ordered-rings}} tactics for automatically solving comparisons on Z (if overflow is allowed, it is an NP-complete problem and human input may be needed: decidability of difference constraints for modular arithmetic (and hence machine integer arithmetic) is NP-complete, by simple reduction from 3-coloring\cite{PointerConstraintsNP}, whereas the same task for full Z is linear in number of variables and constraints).  


\subsection{Separation Logic}

The second problem was related to \textit{pointer aliasing}. These problems are not immediately obvious because C language does not allow us to enforce any memory aliasing restrictions (unlike, say Rust). In formal verification, there is a rigorous model to analyze such kind of problems called \textit{separation logic}. VST provides some automation related to this, however, it is not sufficient, but one can build custom libraries and if restricted to specific domain separation logic proofs can be fully automated. VST supports (undecidable) expressive dependently typed higher-order logic. The built-in tactics can solve simple goals, but much manual proof effort is required to transform the goals. The VST-floyd is based on databases of hints and separation logic lemmmas and tries to apply them in a smart way. We conjecture that the full force of separation logic is unnecessary in most cases and we could restrict ourselves to tractable fragments of separation logic, for which decision procedures has been already implemented within SMT solvers. The latter can be integrated into Coq proofs using SMTCoq\footnote{\url{https://smtcoq.github.io/}}.

\subsection{Specification}

The third issue shows us that your formal verification is only as good as your specification. In this domain we not only have a typical problem of how to validate specification, but also: what kind of specifications make proofs simpler. VST allows to completely separate functional spec from memory spec. However, we noticed that using a C-like spec significantly reduces proof effort and pushes it to the functional level. We experimented with two approaches. In operational semantics proof we used a Coq abstraction of C function and used it as an intermediate specification in the proof. In VST proof we used functional spec written independently from the Coq and used it in the proof of C function correctness directly. This required many lemmas to connect functional spec to C code and complicated the loop invariant. It seems more effective to separate functional aspect of the proof from C-proof to allow for more automation in the proof related to C and then prove the C-like function correctness as a typical functional correctness proof. Moreover, this way we can extract a functional parser and test it.

A nice thing about \texttt{asn1c} compiler is its modular structure, so we can proceed with verification in a modular way and exploit this structure in the proof. For each constructed type its parser will correspond to calling its elements' parsers in a certain order. Hence, we can represent parsers as certain finite state machines that use information provided by the ASN.1 schema to parse the binary input. Alternatively, parsers could be represented using parser combinators mentioned earlier. Decoder functions produced by \texttt{asn1c} compiler use as input a representation of ASN.1 schema, which in Coq we can formalize as a tree.

 \begin{lstlisting}[language=Coq]
Inductive TYPE_descriptor :=
  DEF { tags : list Z ;
        elements : list TYPE_descriptor 
      }.
 \end{lstlisting}

Then a sequence decoder for example will traverse the tree and apply primitive decoders, we can represent it as function that uses nested recursion (as it is implemented in the asn1c code)
 
 \begin{lstlisting}[language=Coq]
 Fixpoint seq_decoder (X : TYPE_descriptor) (ls : list byte)  :=
    match ls with
    | [] =>  OK ls
    | [t] => MORE ls       
    | t :: l :: bs =>          
      if check_tag t (tags X)
        then
          match elements X with 
          | [] => primitive_decoder l bs 
          | XS => (fix elem_seq_decoder XS bs :=
                        match XS with
                        | [] => OK bs
                        | X :: XS =>
                          match seq_decoder X bs with
                            | OK r => elem_seq_decoder XS r 
                            | r => r
                          end
                        end) XS bs
          end
        else ERROR (l::bs)
        end.
 \end{lstlisting}

 or as a FSM (example) or as sequence combinator (example) constructed from the given \texttt{TYPE\_descriptor}. 

\bibliography{paper}
  \bibliographystyle{apalike}


\end{document}
